Impact of grids and dynamical cores in CESM2.2 on the meteorology and climate of the Arctic

### 2021-03-10 arcticgrids.tex
Updates from Overleaf

\title{Impact of grids and dycores in CESM2.2 on the meteorology and climate of the Arctic}

@@

General Circulation Models (GCMs) are a powerful tool for understanding the meteorology and climate of the poles, which are among the most sensitive regions on Earth to global and environmental change. Despite their importance, numerically, not all GCMs treat the polar regions the same way as they do elsewhere in the domain due to the \textit{pole-problem} \cite{W2007JMSJ}. The wide variety of grids structures and numerical techniques used to represent polar regions in conventional GCMs are evaluated in this study for their impacts on the meteorology and climate of the Arctic.

@@

An alternative approach is to use unstructured grids. Unstructured grids permit quasi-uniform grid spacing globally, thereby eliminating the pole-problem. They allow for more flexible grid-structures than latitude-longitude grids, and can support regional grid refinement that may be used to capture higher resolution in the polar regions. Regional refinement of polar regions may be desirable as latitude-longitude grids, by virtue of the convergence of meridians, have higher horizontal resolution in polar regions compared to quasi-uniform grids containing the same degrees of freedom.

The meteorology and climate of the Arctic is charcterized by a range of processes and scales that are difficult to represent in GCMs \cite{BETAL2001MWR,SG2017MWR,VETAL2018TC}. The Arctic geography consists of a polar ocean surrounded by the northern coasts of the northern hemisphere continents and the large island of Greenland. Greenland is covered in a kilometers thick ice sheet, and receives a substantial portion of its snowfall from synoptic-scale storms traversing up the eastern sea-board of North America that rain out over its steep margins. Intense cold season Polar Lows traverse the Arctic's melange of sea-ice and open ocean, inducing gale-speed winds. Large volumes of fresh water drain into the ocean through river outlets in Canada and Russia, iceberg discharge and melt water runoff emanating from the margins of GrIS. As such, the Arctic serves as a test-bed for comparing the simulated meteorology and climate to different numerical treatments of polar regions.

The atmospheric component of the Community Earth System Model, version 2.2 (CESM; \url{http://www.cesm.ucar.edu/models/cesm2/}), the Community Atmosphere Model, version 6.3 (CAM; \url{https://github.com/ESCOMP/CAM/wiki}), supports a diverse number of atmospheric dynamical cores (hereafter referred to as \textit{dycores}). These range from dycores using latitude-longitude grids, the finite-volume \cite<FV;>{L2004MWR} and eulerian spectral transform \cite<EUL;>{CETAL006JC} models, and dycores built on unstructured grids, the spectral-element \cite<SE;>{LetAl2018JAMES} and finite-volume 3 \cite<FV3;>{PL2007JCP} models. The EUL dycore is the oldest dycore in CAM, and is the least supported of all the dycores. FV3 is the newest dycore in CESM, but it was not fully incorporated into the model at the time this work commenced, and so is omitted from this study. The authors instead focus on the two most well supported dycores in CESM, SE and FV. The goal of this study is to characterize the representation of polar regions using the SE and FV dycores, as these models treat the high-latitudes, e.g., the pole-problem, in very different ways.

@@

\section{Methods}\label{sec:methods}
\subsection{Grids}
\subsection{Dynamical cores}
\subsubsection{Finite-volume model}
\subsubsection{Spectral-element model}
\subsection{Physical parameterizations}
\subsection{Experimental desing}
\subsection{Observational datasets}
\subsubsection{ERA5}
\subsubsection{LIVVkit 2.1}
\subsection{TempestExtremes}
\subsection{StormCompositer}

>>>>

\title{Impact of grids and dynamical cores in CESM2.2 on the meteorology and climate of the Arctic}

@@

General Circulation Models (GCMs) are a powerful tool for understanding the meteorology and climate of the high latitudes, which are among the most sensitive regions on Earth to global and environmental change. Despite their importance,  not all GCMs treat the polar regions the same way as they do elsewhere in the domain due to the so called \textit{pole-problem} \cite{W2007JMSJ}. The wide variety of grids structures and numerical techniques used to represent polar regions in conventional GCMs are evaluated in this study for their impacts on the meteorology and climate of the Arctic, with a special focus on the Greenland Ice Sheet.

@@

An alternative approach is to use unstructured grids, which eliminate the pole-problem by adopting a quasi-uniform global grid spacing that includes the poles. These grids allow for more flexible grid-structures than latitude-longitude grids, and can support regional grid refinement that may be used to capture higher resolution in the polar regions. Regional refinement of polar regions may be desirable as latitude-longitude grids, by virtue of the convergence of meridians, have higher horizontal resolution in polar regions compared to quasi-uniform grids containing the same degrees of freedom.

The meteorology and climate of the Arctic is charcterized by a range of processes and scales that are difficult to represent in GCMs \cite{BETAL2001MWR,SG2017MWR,VETAL2018TC}. The Arctic geography consists of a polar ocean surrounded by the northern coasts of the Northern Hemisphere continents and the large island of Greenland. Greenland is covered in a kilometers thick ice sheet, and receives a substantial portion of its snowfall from synoptic-scale storms traversing up the eastern sea-board of North America that rain out over its steep margins. Intense cold season Polar Lows traverse the Arctic's melange of sea-ice and open ocean, inducing gale-speed winds. Large volumes of fresh water drain into the ocean through river outlets in Canada and Russia, and iceberg discharge and melt water runoff emanating from the margins of GrIS. As such, the Arctic serves as a test-bed for comparing the simulated meteorology and climate to different numerical treatments of polar regions.

The goal of this study is to characterize the representation of polar regions using the SE and FV dycores, as these models treat the high-latitudes, e.g., the pole-problem, in very different ways.

The atmospheric component of the Community Earth System Model, version 2.2 (CESM; \url{http://www.cesm.ucar.edu/models/cesm2/}), the Community Atmosphere Model, version 6.3 (CAM; \url{https://github.com/ESCOMP/CAM/wiki}), supports a diverse number of atmospheric dynamical cores (hereafter referred to as \textit{dycores}). These range from dycores using latitude-longitude grids, the finite-volume \cite<FV;>{L2004MWR} and eulerian spectral transform \cite<EUL;>{CETAL006JC} models, and dycores built on unstructured grids, the spectral-element \cite<SE;>{LetAl2018JAMES} and finite-volume 3 \cite<FV3;>{PL2007JCP} models. The EUL dycore is the oldest dycore in CAM, and is the least supported of all the dycores in the current model. FV3 is the newest dycore in CESM, but it was not fully incorporated into the CAM at the time this work commenced, and so is omitted from this study. The authors instead focus on the two most well supported dycores in CESM, SE and FV. %The goal of this study is to characterize the representation of polar regions using the SE and FV dycores, as these models treat the high-latitudes, e.g., the pole-problem, in very different ways.

@@

\section{Methods}\label{sec:methods}
\subsection{Community Earth System Model}
\subsection{Grids}
\subsection{Dynamical cores}
\subsubsection{Finite-volume model}
\subsubsection{Spectral-element model}
\subsection{Physical parameterizations}
\subsection{Experimental desing}
\subsection{Observational datasets}
\subsubsection{ERA5}
\subsubsection{LIVVkit 2.1}
\subsection{TempestExtremes}
\subsection{StormCompositer}
